---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(deckgl)
```

# An R Interface to deck.gl

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/deckgl)](https://cran.r-project.org/package=deckgl) [![Travis-CI Build Status](https://travis-ci.org/crazycapivara/deckgl.svg?branch=master)](https://travis-ci.org/crazycapivara/deckgl) [![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![deck.gl Version](https://img.shields.io/badge/deck.gl-v8.1.0-blue.svg)](https://github.com/uber/deck.gl/releases/tag/v8.1.0)

Deckgl for R makes the open-source JavaScript library [deck.gl](https://deck.gl/) available within R via the [htmlwidgets](https://www.htmlwidgets.org/) package. 

## Installation

```{r cran-installation, eval = FALSE}
install.packages("deckgl")
```

You can install the latest version of deckgl from github with:

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("crazycapivara/deckgl")
```

## Usage

Create a _deckgl_ widget:

```{r create-deckgl-widget, eval = FALSE}
deckgl()
```

Add a base map:

```{r basemap, eval = FALSE}
deckgl() %>%
  add_basemap()
```

Add any kind of layers:

```{r add-layers, eval = FALSE}
# Grid layer example
data("sf_bike_parking")

props <- list(
  extruded = TRUE,
  cellSize = 200,
  elevationScale = 4,
  getPosition = ~lng + lat,
  tooltip = "Count: {{count}}"
)

deckgl(zoom = 11, pitch = 45) %>%
  add_basemap() %>%
  add_grid_layer(data = sf_bike_parking, properties = props)
```

## Layers

Due to the generic function `add_layer` any kind of layer defined in the [deckgl-api-reference](https://deck.gl/#/documentation/deckgl-api-reference) is supported. The type of layer is chosen via the `class_name` parameter, e. g. `ScatterplotLayer` or `GeoJsonLayer`. Usually you will not use the generic function but one of the `add_*_layer` shortcut functions instead:

```{r add-layers-generic-vs-shortcut, eval = FALSE}
# Generic function
deckgl() %>%
  add_layer("ArcLayer", "arc-layer", data, properties)

# Shortcut function
deckgl() %>%
  add_arc_layer("arc-layer", data, properties)
```

## Data

The `data` parameter can either be an __url__ to fetch data from or a __data object__.

## Layer Props

The `properties` parameter is a __named list__ with names corresponding to the __properties__ defined in the [deckgl-api-reference](https://deck.gl/#/documentation/deckgl-api-reference) for the given __layer__ type (class). For the example above see the [grid-layer-api-reference](https://deck.gl/#/documentation/deckgl-api-reference/layers/grid-layer).
In case the property is a __data accessor__ (usually a function in the form of a lambda expression) just use the `JS` function in R to pass it to the layer object. In the example above this is done for the properties `getPosition` and `getTooltip`. It is also possible to pass properties as named arguments via the `...` parameter. They are appended to the properties list. Same properties will be overwritten.

## Tooltips

The tooltip for a layer can be set via the `tooltip` parameter. You can either pass a single template
string or a list with the following properties (see also `use_tooltip`):

* `html`: The `innerHTML` of the tooltip.
* `style`: A `cssText` string that will modefiy the default style of the tooltip.

### Tooltip template Syntax

The tooltip string is a _mustache_ template in which variable names are identified
by the double curly brackets (_mustache_ tags) that surround them. The variable names available to the
template are given by deck.gl’s `pickingInfo.object` and vary by layer.

See [mustache](https://github.com/janl/mustache.js) for a complete syntax overwiew.

## Run Examples

You can run the [api-examples](https://github.com/crazycapivara/deckgl/tree/master/inst/examples/deckgl-api-reference) from the `add_*_layer` functions with `example(add_*_layer)`. For the `IconLayer` it looks like this:

```{r run-examples, eval = FALSE}
# Sys.setenv(MAPBOX_API_TOKEN = "yourSuperSecretApiToken")
example(add_icon_layer)
```

## Development

The JavaScript library of the _deckgl widget_ uses [webpack](https://webpack.js.org/) as module bundler.
Therefore, you need [node.js](https://nodejs.org) to build the _widget_ library.

First of all you need to clone the repository:

```bash
git clone https://github.com/crazycapivara/deckgl.git
```

Then you can build the library from inside the `javascript` folder like this:

```bash
cd deckgl/javascript

npm run build
```

All JavaScript code is located in the `src` folder and test components go to the `src/test-components` folder.

To test the library just run the following command. This will spin up the [webpack-dev-server](https://github.com/webpack/webpack-dev-server):

```bash
npm run start
```

## Concept

Deckgl for R stays as close as possible to the JavaScript API so that usually all parameters of its JavaScript pendants are supported. Therefore, you need to check the [deckgl-api-reference](https://deck.gl/#/documentation/deckgl-api-reference) of the JavaScript framework to get information about the parameters you can pass to the R-functions mostly as named lists or named arguments (`...` parameter). Use the `JS` function if you need to pass any kind of JavaScript code, as it is the case for __data accessors__.
 
[GridLayer](https://deck.gl/#/documentation/deckgl-api-reference/layers/grid-layer) API-example:
 
```JavaScript
// JavaScript code

const layer = new GridLayer({
  id: "grid-layer",
  data: data,
  extruded: true,
  cellSize: 200,
  elevationScale: 4,
  getPosition: d => d.COORDINATES,
});
```

```{r eval = FALSE}
# Corresponding R code

deck <- deckgl() %>%
  add_grid_layer(
    id = "grid-layer",
    data = data,
    extruded = TRUE,
    cellSize = 200,
    elevationScale = 4,
    getPosition = JS("d => d.COORDINATES")
  )
```

In this example all properties are passed as named arguments.

You can test your R code like this:

```{r}
properties <- list(
  extruded = TRUE,
  cellSize = 200,
  elevationScale = 4,
  getPosition = get_property("COORDINATES")
)

htmlwidgets:::toJSON2(properties, pretty = TRUE)
```

## Documentation

- [deckgl for R](https://crazycapivara.github.io/deckgl/) (pkgdown-site)
- [example scripts](https://github.com/crazycapivara/deckgl/tree/master/inst/examples)
- [deckgl-api-reference](https://deck.gl/#/documentation/deckgl-api-reference) (JavaScript framework)

## Notes

* It is a known issue that the deckgl widget might not be visible in the viewer pane of RStudio. Just open it in your browser by clicking __Show in new window__ and everything will be fine.
* The [documentation](https://crazycapivara.github.io/deckgl/) is work in progress. Please check also the [examples](https://github.com/crazycapivara/deckgl/tree/master/inst/examples) as a starting point.
