---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# An R Interface to deck.gl

Deckgl for R makes the open-source JavaScript library [deck.gl](https://deck.gl/) available within R via the [htmlwidgets](https://www.htmlwidgets.org/) package. 

## Notes

- It is a known issue that the deckgl widget might not be visible in the viewer pane of RStudio. Just open it in your browser by clicking _Show in new window_ and everything will be fine.
- The documentation is work in progress. Please check the [examples](inst/examples) as a starting point.
- You do _not_ need a [mapbox](https://www.mapbox.com/) api key to use this package. It is always optional to add a base map from mapbox to your widget.

## Installation

You can install deckgl from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("crazycapivara/deckgl")
```

## Quickstart

Just check if everything is fine:

```{r does-it-work, eval = FALSE}
library(deckgl)

does_it_work()
#> You should see a text layer telling you that it works.

# Or in case you do have an api token from mapbox ...
does_it_work("yourSuperSecretApiToken")
#> Output should be the same as above but rendered on top of a base map from mapbox.
```

Show some sample data:

```{r sample-data, eval = FALSE}
# Pass 'data = NULL' to load some sample data for the hexagon layer
deckgl() %>%
  add_hexagon_layer(data = NULL) %>%
  add_mapbox_basemap("yourSuperSecretApiToken") # optional
```

## Usage

First you need to create a deckgl widget:

```{r create-deckgl-widget, eval = FALSE}
deckgl()
```

Add a base map from mapbox (optional):

```{r basemap, eval = FALSE}
token <- "yourSuperSecretApiToken"

deckgl() %>%
  add_mapbox_basemap(token, style = "mapbox://styles/mapbox/dark-v9")
```

Add any kind of layers:

```{r add-layers, eval = FALSE}
# Grid layer example

sample_data <- "https://raw.githubusercontent.com/uber-common/deck.gl-data/master/website/sf-bike-parking.json"

properties <- list(
  pickable = TRUE,
  extruded = TRUE,
  cellSize = 200,
  elevationScale = 4,
  getPosition = JS("data => data.COORDINATES"),
  onHover = JS("({object}) => { if(object) console.log(object.count) }")
)

deckgl(zoom = 11, pitch = 45) %>%
  add_grid_layer(data = sample_data, properties = properties) %>%
  add_mapbox_basemap(token) # optional
```

- The `data` parameter can either be an _url_ to fetch data from or an _R data object_
- The `properties` parameter is a _named list_ with names corresponding to the properties defined in the [deckgl-api-reference](https://deck.gl/#/documentation/deckgl-api-reference) for the given layer class. For the example above see [api-reference grid-layer](https://deck.gl/#/documentation/deckgl-api-reference/layers/grid-layer).
In case the property is a function (usually a lambda expression), just use the `JS` function in R to pass it to the layer object. In the example above this done for the properties `getPosition` and `onHover`.

## Layers

Due to the generic function `add_layer` any kind of layer defined in the deckgl-api-reference is supported. The type of layer is chosen via the `class_name` parameter, e. g. `ScatterplotLayer` or `GeoJsonLayer`. Usually you will not use the generic function but one of the `add_*_layer` shortcut functions instead:

```{r add-layers-generic-vs-shortcut, eval = FALSE}
# Generic function
deckgl() %>% add_layer("ArcLayer", "arc-layer", data, properties)

# Shortcut function
deckgl() %>% add_arc_layer("arc-layer", data, properties)
```
